---
title: "File Enumerator"
author: Lukas
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(lese)
library(data.table)
library(ggplot2)
library(dplyr)
library(kableExtra)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

lecture_files <- rbindlist(lapply(lectures(), \(lecture) {
  files <- fs::dir_ls(lecture, all = TRUE, recurse = TRUE, type = "file")
  # Remove first element (lecture name) from path
  files <- fs::path_rel(files, start = lecture)

  # Filter out .git
  files <- fs::path_filter(files, glob = ".git/*", invert = TRUE)
  fs::path_split(files[5:10])

  data.table(
    lecture = lecture,
    file = fs::path_file(files),
    parent_dir = fs::path_dir(files),
    extension = tolower(fs::path_ext(files))
  )
}))


enumerate_files <- function(lecture) {
  dir <- fs::dir_ls(lecture, recurse = TRUE, type = "dir")
  # Remove first element (lecture name) from path

  file_list <- lapply(dir, \(d) {
    dt <- data.table(
      dir = fs::path_rel(d, start = lecture),
      file = fs::dir_ls(path = d, recurse = TRUE, type = "file") |>
        fs::path_file()
    )
    if (nrow(dt) == 0) {
      return(data.table())
    }
    dt
  })
  file_dt <- as.data.table(dplyr::bind_rows(file_list))

  file_dt[, depth := lengths(fs::path_split(dir))]
  file_dt[, extension := tolower(fs::path_ext(file))][]
}

count_files_per_dir <- function(lecture) {
  file_tbl <- enumerate_files(lecture)

  file_tbl |>
    count(dir, depth, sort = TRUE) |>
    arrange(desc(n)) |>
    reactable::reactable(
      sortable = TRUE,
      filterable = TRUE,
      defaultPageSize = 20
    )
}

count_extensions_per_dir <- function(lecture) {
  file_tbl <- enumerate_files(lecture)

  file_tbl |>
    count(dir, depth, extension, sort = TRUE) |>
    arrange(desc(n)) |>
    filter(n > 30) |>
    reactable::reactable(
      sortable = TRUE,
      filterable = TRUE,
      defaultPageSize = 20
    )
}

total_counts <- function(lecture) {
  lecture_files |>
    filter(.data$lecture == .env$lecture) |>
    count(lecture, sort = TRUE)
}

total_counts_ext <- function(lecture) {
  lecture_files |>
    filter(.data$lecture == .env$lecture) |>
    count(extension, sort = TRUE) |>
    filter(n >= 30)
}
```

## lecture_i2ml

```{r}
total_counts("lecture_i2ml")
total_counts_ext("lecture_i2ml")
```


```{r}
count_files_per_dir("lecture_i2ml")
```

Counting file extension counts, only showing those with counts > 30

```{r}
count_extensions_per_dir("lecture_i2ml")
```

## lecture_sl

```{r}
total_counts("lecture_sl")
total_counts_ext("lecture_sl")
```

```{r}
count_files_per_dir("lecture_sl")
```

Counting file extension counts, only showing those with counts > 30

```{r}
count_extensions_per_dir("lecture_sl")
```

## lecture_advml

```{r}
total_counts("lecture_advml")
total_counts_ext("lecture_advml")
```

```{r}
count_files_per_dir("lecture_advml")
```

Counting file extension counts, only showing those with counts > 30

```{r}
count_extensions_per_dir("lecture_advml")
```

## lecture_optimization

```{r}
total_counts("lecture_optimization")
total_counts_ext("lecture_optimization")
```


```{r}
count_files_per_dir("lecture_optimization")
```

Counting file extension counts, only showing those with counts > 30

```{r}
count_extensions_per_dir("lecture_optimization")
```


## lecture_iml

```{r}
total_counts("lecture_iml")
total_counts_ext("lecture_iml")
```

```{r}
count_files_per_dir("lecture_iml")
```

Counting file extension counts, only showing those with counts > 30

```{r}
count_extensions_per_dir("lecture_iml")
```
